---
title: "Data Appendix"
author: "Kiersten Hamby,Breanna Ranglall,Krutika Tekwani"
format: pdf
editor: visual
execute: 
  echo: true
  self-contained: true
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(skimr)
```

## Data Appendix

## Group Members

Kiersten Hamby, Breanna Ranglall, Krutika Tekwani

## Loading Data In

```{r}
Athletics=read.csv("~/OneDrive - University of St. Thomas/STAT320/GroupE2AthleticsData/GroupE2-AthleticsSpendingData.csv")
```

## Skim

```{r}
Athletics%>%
  select(institution_name,state_cd,classification_name,EFMaleCount,EFFemaleCount,EFTotalCount,sector_name,STUDENTAID_MEN,STUDENTAID_WOMEN,STUDENTAID_TOTAL,HDCOACH_SALARY_MEN,HDCOACH_SALARY_WOMEN,RECRUITEXP_MEN,RECRUITEXP_WOMEN,RECRUITEXP_TOTAL,NUM_HDCOACH_MEN,NUM_HDCOACH_WOMEN)%>%
  skim()
```

## Variable Names/Types

institution_name

state_cd

classification_name

EFMaleCount

EFFemaleCount

EFTotalCount

sector_name

STUDENTAID_MEN: 13 missing values

STUDENTAID_WOMEN: 31 missing values

STUDENTAID_TOTAL

HDCOACH_SALARY_MEN: 14 missing values

HDCOACH_SALARY_WOMEN: 32 missing values

RECRUITEXP_MEN: 13 missing values

RECRUITEXP_WOMEN: 31 missing values

RECRUITEXP_TOTAL

NUM_HDCOACH_MEN: 15 missing values

NUM_HDCOACH_WOMEN: 33 missing values

The variable names are readable and clear. One point of confusion is what 'EF' means in the count of students. Sector name refers to the public/private status of a school. It is convenient that the data already specifies men's, women's, and totals for the variables. Some variables include underscores in the name.

All variables except those discussed below, Categorical, are numeric variables which is what they should be.

## Category Names

institution_name

state_cd

classification_name

sector_name

All non-binary categorical variables. Clearly named, though they do include underscores. These are all classified as character variables which is what we would expect.

## Max and Min for Variables

EFMaleCount: Max=34,767, Min=1

EFFemaleCount: Max=29,793, Min=4

EFTotalCount: Max=64,560, Min=62

STUDENTAID_MEN: Max=15,953,475, Min=0

STUDENTAID_WOMEN: Max=14,979,930, Min=0

STUDENTAID_TOTAL: Max=28,909,727, Min=0

HDCOACH_SALARY_MEN: Max=2,387,189, Min=0

HDCOACH_SALARY_WOMEN: Max=601,582, Min=0

RECRUITEXP_MEN: Max=5,331,795, Min=0

RECRUITEXP_WOMEN: Max=962,902, Min=0

RECRUITEXP_TOTAL: Max=6,294,697, Min=0

NUM_HDCOACH_MEN: Max=106, Min=1

NUM_HDCOACH_WOMEN: Max=101, Min=2

The only point of concern is that so many variables have a minimum of 0 and we would not expect to see a university with no recruiting expenses, for example. We will look into what is affecting this, potentially the sector of the school.

## Data Wrangling, Data Cleaning

-Only have the parent school if there are multiple campuses to maintain independence

-Deleting unused variables as there are more than 4,000 variables

-Look into schools that have 0 as a minimum for expenses/student aid/etc. to see if we can find a cause or need to remove these observations from our data.
